Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.Rmd
more changes on the plot for the last simulation and writing

```{r noisy-better-geo, fig.env = "widefigure", fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index. The trace plot shows the failure of pseudo derivative on non-smooth index. The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index."}
```

@@

```{r kol-result, fig.height = 6, fig.cap="One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better}, starting from 20 different seeds. "}
```

The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present. The dataset used is \code{boa6} where `x2` and `x7` are informative. The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively. Hence, the global maximum happens when variable `x7` is found.

<!-- Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \ref{fig:kol-better}) and `search_better_random` (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}). Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.  The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths. -->

The conclusion from this experiment is that the usage of `search_better_random` and increasing the search space are methods that can avoid getting trapped in local maxima but the solution will still depend on the starting points of the simulation and the seed used.

>>>>

```{r noisy-better-geo, fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index. The trace plot shows the failure of pseudo derivative on non-smooth index. The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index."}
```

@@

```{r kol-result, fig.height = 6, fig.cap="One-D projection of data \\code{boa6} on \\code{norm\\_kol} index,  optimised by \\code{search\\_better} and \\code{search\\_better\\_random}. Two different neighbourhood size of 0.5 and 0.7 is used. The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The color of the path is based on whether the optimiser finds the global or local optimum, after polishing."}
```

The next experiment compares two optimisers: `search_better` and `search_better_random` for 20 simulation paths. Two search neighbourhood sizes, `alpha`, of 0.5 and 0.7 are also compared to understand the effect of neighbourhood size where a larger value indicating a wider search. Figure \ref{fig:kol-result} shows 80 (20 * 4) PCA projected interpolation paths faceted by optimiser and alpha. Several results can be drawn through the comparison: 1) With `search_better` and `alpha = 0.5`, despite being the simplest and fastest, the optimiser fails to optimise for three instances where they finish neither near the local nor the global optimum. 2) With the same optimiser but a larger neighbourhood size `alpha = 0.7`, more seeds have found the global optimum. 3) Comparing different optimisers `search_better` and `search_better_random` with the same search size of 0.5, `search_better_random` finishes closer to the theoretical for the ones that find the local optimum. This indicates the optimiser is working harder to examine if there is other optimum in the space but a small search neighbourhood size makes it hard to conduct a holistic search of the whole space. 4) With `search_better_random` and a larger neighbourhood size of 0.7, more seeds have found the global optimum. Also some points are getting closer to the position of the theoretical best, as comparing with `alpha = 0.5`, and this would indicate fewer work for the subsequent polish.

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.tex
more changes on the plot for the last simulation and writing

\begin{Schunk}
\begin{widefigure}

@@

\caption[1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index]{1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index. The trace plot shows the failure of pseudo derivative on non-smooth index. The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
\end{widefigure}
\end{Schunk}

@@

\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds. }\label{fig:kol-result}
\end{figure}
\end{Schunk}

The second experiment with the noisy index is to understand how the
optimisers perform when a local maximum is present. The dataset used is
\code{boa6} where \texttt{x2} and \texttt{x7} are informative. The two
theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0,
0{]} with index value 0.176 and 0.235, respectively. Hence, the global
maximum happens when variable \texttt{x7} is found.

The conclusion from this experiment is that the usage of
\texttt{search\_better\_random} and increasing the search space are
methods that can avoid getting trapped in local maxima but the solution
will still depend on the starting points of the simulation and the seed
used.

>>>>

@@

\caption[1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index]{1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index. The trace plot shows the failure of pseudo derivative on non-smooth index. The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
\end{figure}
\end{Schunk}

@@

\caption[One-D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by \code{search\_better} and \code{search\_better\_random}]{One-D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by \code{search\_better} and \code{search\_better\_random}. Two different neighbourhood size of 0.5 and 0.7 is used. The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The color of the path is based on whether the optimiser finds the global or local optimum, after polishing.}\label{fig:kol-result}
\end{figure}
\end{Schunk}

The next experiment compares two optimisers: \texttt{search\_better} and
\texttt{search\_better\_random} for 20 simulation paths. Two search
neighbourhood sizes, \texttt{alpha}, of 0.5 and 0.7 are also compared to
understand the effect of neighbourhood size where a larger value
indicating a wider search. Figure \ref{fig:kol-result} shows 80 (20 * 4)
PCA projected interpolation paths faceted by optimiser and alpha.
Several results can be drawn through the comparison: 1) With
\texttt{search\_better} and \texttt{alpha\ =\ 0.5}, despite being the
simplest and fastest, the optimiser fails to optimise for three
instances where they finish neither near the local nor the global
optimum. 2) With the same optimiser but a larger neighbourhood size
\texttt{alpha\ =\ 0.7}, more seeds have found the global optimum. 3)
Comparing different optimisers \texttt{search\_better} and
\texttt{search\_better\_random} with the same search size of 0.5,
\texttt{search\_better\_random} finishes closer to the theoretical for
the ones that find the local optimum. This indicates the optimiser is
working harder to examine if there is other optimum in the space but a
small search neighbourhood size makes it hard to conduct a holistic
search of the whole space. 4) With \texttt{search\_better\_random} and a
larger neighbourhood size of 0.7, more seeds have found the global
optimum. Also some points are getting closer to the position of the
theoretical best, as comparing with \texttt{alpha\ =\ 0.5}, and this
would indicate fewer work for the subsequent polish.
